\documentclass[12pt,letterpaper]{article}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{fancyhdr}

% Set up the headers and footers
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[C]{Information Disclosure Statement Under 37 CFR §1.97 and §1.98}
\fancyfoot[C]{Page \thepage}

% Define a title command
\newcommand{\papertitle}[1]{\noindent\begin{center}\textbf{#1}\end{center}}

\begin{document}

\begin{center}
\textbf{\large INFORMATION DISCLOSURE STATEMENT BY APPLICANT}\\
\vspace{0.25in}
\textbf{\large 37 CFR §1.97 and §1.98}\\
\vspace{0.5in}
\end{center}

\begin{tabular}{ll}
Application Number: & 19/169,399 \\
Filing Date: & April 3, 2025 \\
First Named Inventor: & Luis Minier \\
Art Unit: & [Pending Assignment] \\
Examiner Name: & [Pending Assignment] \\
Title: & A Hybrid Computational Framework for Quantum and Resonance Simulation \\
\end{tabular}

\vspace{0.5in}

\noindent I hereby certify that each item of information contained in this Information Disclosure Statement was first cited in any communication from a foreign patent office in a counterpart foreign application not more than three months prior to the filing of this statement.

\vspace{0.25in}

\noindent Applicant submits herewith the following information disclosure statement, including Form PTO/SB/08a, for consideration by the Examiner. The publications are submitted to comply with the duty of disclosure under 37 CFR §1.56.

\vspace{0.25in}

\papertitle{1. Publication Details}

\begin{longtable}{|p{1.5in}|p{4.5in}|}
\hline
\textbf{Publication Type} & Non-Patent Literature \\
\hline
\textbf{Title} & QuantoniumOS V3: Advanced Testing Suite for Symbolic Resonance Encryption (Version 4) \\
\hline
\textbf{Author(s)} & Luis, M. \\
\hline
\textbf{Publication Date} & 2024 \\
\hline
\textbf{Publisher} & Zenodo \\
\hline
\textbf{DOI} & \url{https://doi.org/10.5281/zenodo.15256826} \\
\hline
\textbf{Pages} & 1-87 \\
\hline
\textbf{Relevance} & This publication contains implementation details, test results, and technical specifications that demonstrate the enablement and working implementation of the claimed invention. \\
\hline
\end{longtable}

\newpage
\papertitle{2. Additional Publications}

\begin{longtable}{|p{1.5in}|p{4.5in}|}
\hline
\textbf{Publication Type} & Non-Patent Literature \\
\hline
\textbf{Title} & Container Architecture for Quantum-Resistant Storage \\
\hline
\textbf{Author(s)} & Luis, M. \\
\hline
\textbf{Publication Date} & 2023 \\
\hline
\textbf{Publisher} & Proceedings of the International Symposium on Quantum Information Processing \\
\hline
\textbf{Pages} & 342-357 \\
\hline
\textbf{Relevance} & This publication contains architectural details relevant to the container validation aspects of the claimed invention. \\
\hline
\end{longtable}

\begin{longtable}{|p{1.5in}|p{4.5in}|}
\hline
\textbf{Publication Type} & Non-Patent Literature \\
\hline
\textbf{Title} & Human-Verifiable Cryptography through Waveform Visualization \\
\hline
\textbf{Author(s)} & Luis, M. \\
\hline
\textbf{Publication Date} & 2024 \\
\hline
\textbf{Publisher} & Journal of Applied Cryptography \\
\hline
\textbf{Volume/Issue} & Vol. 28, Issue 4 \\
\hline
\textbf{Pages} & 112-128 \\
\hline
\textbf{Relevance} & This publication discusses visualization techniques for cryptographic waveforms relevant to the user interface aspects of the claimed invention. \\
\hline
\end{longtable}

\newpage
\papertitle{3. Mapping of Publications to Patent Claims}

\noindent The following table maps specific sections of the submitted publications to the claims in this patent application, demonstrating the enablement of each claim:

\begin{longtable}{|p{1in}|p{2.5in}|p{2.5in}|}
\hline
\textbf{Claim Number} & \textbf{Claim Summary} & \textbf{Publication Sections Supporting Enablement} \\
\hline
\textbf{Claim 1} & 
Resonance Fourier Transform with bidirectional mapping between waveform data and frequency domain, allowing perfect reconstruction with cryptographic properties & 
\begin{itemize}
\item \textbf{Zenodo Paper}: 
  \begin{itemize}
  \item Section 3.1 (pp. 12-18): "Resonance Fourier Transform Algorithm"
  \item Section 4.2 (pp. 24-29): "Bidirectional Transformation Testing"
  \item Appendix A (pp. 72-74): "RFT Implementation Code"
  \end{itemize}
\item \textbf{Applied Cryptography Paper}: 
  \begin{itemize}
  \item Section 2 (pp. 114-116): "Waveform Properties in Cryptographic Applications"
  \end{itemize}
\end{itemize} \\
\hline
\textbf{Claim 2} & 
Geometric waveform hashing for secure container validation where hash values function as both identifiers and keys & 
\begin{itemize}
\item \textbf{Zenodo Paper}:
  \begin{itemize}
  \item Section 3.3 (pp. 19-23): "Geometric Waveform Hashing"
  \item Section 5.1 (pp. 34-38): "Container Validation Process"
  \item Appendix B (pp. 75-77): "Hash Generation Algorithm"
  \end{itemize}
\item \textbf{Container Architecture Paper}:
  \begin{itemize}
  \item Section 3 (pp. 345-349): "Hash-Key Duality in Container Access"
  \item Section 4 (pp. 350-352): "Wave Coherence Verification"
  \end{itemize}
\end{itemize} \\
\hline
\textbf{Claim 3} & 
Symbolic character variables for encryption operations that function as addressable numeric units within a computational framework & 
\begin{itemize}
\item \textbf{Zenodo Paper}:
  \begin{itemize}
  \item Section 3.2 (pp. 18-19): "Symbolic Character Variables"
  \item Section 4.3 (pp. 30-33): "Vectorized Operations on Symbolic Units"
  \item Appendix C (pp. 78-80): "Wave Primitives Implementation"
  \end{itemize}
\end{itemize} \\
\hline
\textbf{Claim 4} & 
Quantum simulation with secure algorithm protection supporting up to 150 qubits with strict separation between visual interface and core algorithms & 
\begin{itemize}
\item \textbf{Zenodo Paper}:
  \begin{itemize}
  \item Section 3.4 (pp. 23-24): "Quantum Simulation Module"
  \item Section 5.3 (pp. 42-46): "Security Architecture for Algorithm Protection"
  \item Section 6.2 (pp. 51-54): "150-Qubit Performance Testing"
  \item Appendix D (pp. 81-83): "API Interface Design"
  \end{itemize}
\item \textbf{Applied Cryptography Paper}:
  \begin{itemize}
  \item Section 4 (pp. 120-123): "Interface-Algorithm Separation Patterns"
  \end{itemize}
\end{itemize} \\
\hline
\end{longtable}

\newpage
\papertitle{4. Performance Test Results from Publication}

\noindent The following table, extracted from the submitted Zenodo publication (p. 52), demonstrates performance testing results for key components of the claimed invention:

\begin{longtable}{|p{2.5in}|p{1in}|p{1in}|}
\hline
\textbf{Operation} & \textbf{Time (ms)} & \textbf{Memory (MB)} \\
\hline
RFT (32-point) & 2.3 & 1.5 \\
\hline
IRFT (32-point) & 2.5 & 1.5 \\
\hline
Container Sealing & 5.8 & 3.2 \\
\hline
Container Validation & 4.2 & 2.8 \\
\hline
Quantum Circuit (10 qubits) & 12.7 & 15.6 \\
\hline
Quantum Circuit (50 qubits) & 456.3 & 128.4 \\
\hline
\end{longtable}

\vspace{0.5in}

\papertitle{5. System Architecture Diagram from Publication}

\noindent The following architecture diagram, extracted from the submitted Zenodo publication (p. 15), illustrates the implementation of the layered system architecture supporting all claimed functionality:

\begin{center}
\begin{tabular}{|p{1.5in}|p{4in}|}
\hline
\textbf{Presentation Layer} & Client Applications (Web \& Desktop) \\
\hline
\textbf{Security Layer} & API Layer \& Security Middleware \\
\hline
\multirow{3}{*}{\textbf{Application Layer}} & 
\begin{tabular}{|p{1.2in}|p{1.2in}|p{1.2in}|}
\hline
Encryption \& Hashing Module & Core Engine \& Container Orchestration & Quantum Simulation Module \\
\hline
\end{tabular} \\
\hline
\textbf{Infrastructure Layer} & Infrastructure (DB, Secrets, Rate-Limit) \\
\hline
\end{tabular}
\end{center}

\vspace{0.5in}

\noindent I certify that the information cited in this communication complies with the duty of disclosure under 37 C.F.R. § 1.56 and is relevant to the examination of this application.

\vspace{1in}

\begin{tabular}{ll}
Signature: & \underline{\hspace{3in}} \\
\\
Printed Name: & Luis Minier \\
\\
Date: & \underline{\hspace{3in}} \\
\end{tabular}

\end{document}